Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-12-04 paper/main.tex
add some changes suggested by christian

We model the force exerted by the fluid flow onto the individual vertices by
\begin{align}
    \vec{G}_i &= C
    \underbrace{|\vec{x}_i - \vec{x}_{i+1}|\sin(\beta_i)}_\text{proj. rod segment length}
    \underbrace{(yR - y^2)}_\text{flow contribution}\vec{e}_x\\
    \vec{F}_{i,\text{external}} &= \frac{1}{2}(\vec{G}_{i-1}+\vec{G}_i)
\end{align}
which is proportional to the length of the rod segment projected along the y-axis.
Similarly to Stoke's Law~\cite{Stokes1856}, it acts along the axis of the fluid flow (x-axis).
Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
The flow profile for this particular setup is proportional~\cite{Wolschin2021} to $(2yR-y^2)$ where
$y=0.5(\vec{x}_i+\vec{x}_{i+1})_y-y_b$ is the vertical displacement measured from the end of the
tube $y_b$ (see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber
along the y-axis which is hidden in Figure~\ref{fig:amir-bending-simulation} (A).
In particular, this enforces the correct boundary values the flow profile and the resulting force
$\vec{F}_{i,\text{external}}|_{y=0}=0$.
In addition we confine vertices who belong to segments that are partially within the tube
$(\vec{x}_i)_y\leq y_b$ via $(\vec{x}_{i+1})_x=x_0$ and also fix the value for the first vertex
$\vec{x}_0$ such that the bacterium is attached to the bottom.

>>>>

The flow profile for this setup is proportional~\cite{Wolschin2021} to $(2y_iy_\text{max}-y_i^2)$
where $y_i$ is the vertical displacement measured from the end of the tube and $y_\text{max}$
is the total size of the champer along the y-axis which is hidden in
Figure~\ref{fig:amir-bending-simulation} (A).
For every segment that connects two vertices $\vec{x}_i,\vec{x}_{i+1}$, we evaluate the flow profile
at $y_i=0.5(\vec{x}_i+\vec{x}_{i+1})_y$.
In particular, this enforces the correct boundary values the flow profile and the resulting force
$\vec{F}_{i,\text{external}}|_{y=0}=0$.
We therefore model the force exerted by the fluid flow onto the individual vertices by
\begin{align}
    \vec{G}_i &= C
    \underbrace{|\vec{x}_i - \vec{x}_{i+1}|\sin(\beta_i)}_\text{proj. rod segment length}
    \underbrace{(y_iy_\text{max} - y_i^2)}_\text{flow contribution}\vec{e}_x\\
    \vec{F}_{i,\text{external}} &= \frac{1}{2}(\vec{G}_{i-1}+\vec{G}_i)
\end{align}
which is proportional to the length of the rod segment projected along the y-axis.
Similarly to Stoke's Law~\cite{Stokes1856}, it acts along the axis of the fluid flow (x-axis).
Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
In addition we confine vertices who belong to segments that are within the tube
via $(\vec{x}_{i+1})_x=x_0$ and fix the value for the first vertex
$\vec{x}_0$ such that the bacterium is attached to the bottom.
